### 2019-01-22 tex_v2/manuscript.tex
intro 1st pass

@@

	%title and author block
	{\title{Microstructure and mechanical behavior of the acorn weevil exoskeleton}

	\date{\today}

@@

%introduction - nature materials requires that introduction is given NO SECTION HEADER
%Brief explanation of relevant behavior and biology, prior work on ecology, head shape, cuticle mechanics, and rostral microstructure.

%~650 words here

	We report novel modifications to the composite microstructure of the exoskeleton in the snout of acorn weevils (Coleoptera: Curculionidae) belonging to the genus \textit{Curculio} Linnaeus, 1756.

	As a weevil (snout beetle), members of the genus \textit{Curculio} are typified by the presence of a highly elongate structure on the head, called the rostrum (snout). 
	This structure is a hollow, strongly curved (over 90$\degree$), cylindrical extension of the exoskeleton of the otherwise nearly-spherical head, which bears at its apex the terminal chewing mouthparts. 
	The space inside of the rostrum contains the esophagus, various muscles and tendons used for feeding, and hemolymph that serves as a rough equivalent to blood in insects.
	By contrast, the solid shell of the rostrum is comprised entirely of cuticle, which can be considered a laminate composite consisting of various arrangements of chitin fibers embedded in a protein matrix (see \cref{sec:microstructure}).
	Acorn weevils use this structure to excavate sites for egg-laying (oviposition) and feeding on a variety of fruits, including acorns, Japanese camellia, hazelnuts, pecans, chestnuts, and chinquapins.

	During oviposition, a female engages in a unique "drilling" behavior that causes significant, apparently elastic, deformation of the rostrum.
	The female will insert the snout into an incision made with the mandibles, eating the material as she proceeds, while rotating her head and body around the perimeter of the bore-hole.
	Once the apex of the snout is fully inserted, she will push up and forward with her front legs, forcing the rostrum to bend until it is nearly straight.
	The female will maintain tension on the rostrum in this position, continuing to ingest the substrate and rotate around the bore-hole, while slowly inserting the rostrum further into the excavated channel.
	Once the rostrum is fully inserted, usually up to the eyes, she will pull her snout from the bore-hole and deposit several eggs into the site.
	By maintaining constant tension on the rostrum and rotating around the bore-hole, the female is able to flex the snout into a near-perfectly straight configuration and thereby produce a linear channel into the fruit.
	While this behavior has been observed in many species of \textit{Curculio}, we have lacked a fundamental understanding of how female \textit{Curculio} rostra can withstand the repeated, often extreme bending incurred during the process of oviposition.

	We have found that the composite profile of the rostrum is strongly differentiated from the head capsule and other body parts, with modification of both the relative layer thicknesses and fiber orientation angles of cuticle regions (viz. exocuticle and endocuticle), which we describe in detail below.
	We posit that these modifications enable the snout to be flexed until straight while remaining within the elastic limits of the material, mitigating the risk of structural damage, and without evident alteration of the mechanical properties of the individual components of the cuticle across the structure and between species.
	Thus, the flexibility and tensile strength of the rostrum appear to be derived \emph{exclusively} from modification of the composite architecture of the exoskeleton.

	Support for this hypothesis has come from three lines of evidence:

	\begin{enumerate}
	\item Examination of the cuticle microstructure across the length of the snout has revealed consistent modification to the composite structure of the rostrum among \textit{Curculio} species.
	\item Tensile testing of the rostrum has demonstrated that the mechanical strength of the cuticle components are consistent along the length of the structure and between species.
	\item Fatigue testing has shown that a highly curved rostrum is capable of flexing hundreds of thousands of times without damage to the structure, and is apparently elastic.
	\end{enumerate}

	We additionally describe the fracture mechanics of the snout, as pertains to both cuticle composite structure and tensile behavior, and consider how modification of the cuticle may reduce the risk of rostral fracture during oviposition.
	To our knowledge, this is the first time that a modified composite profile has been reported as a means of enhancing structural elasticity in the insect exoskeleton.

	MIGHT BE GOOD TO SIGNPOST HERE... we discuss each of these lines of evidence in the sections below beginning with a consideration of the impact of microstructure, then we talk about mechanical testing, blah blah. Emphasize that this is key to predicting and understanding the mechanical behavior of the snout during bending.

%stick to 600 words for each results section and 500 words for conclusions
	\section{Microstructure of the rostrum}\label{sec:microstructure}
	%~700 words here (1500 total, shorten later)
		Insect cuticle is made up of chitin and numerous uncharacterized proteins, which effectively act as a matrix into which the chitin is embedded \cite{Nikolov2011,Nikolov2010,Vincent2004}.
		The arrangement of the embedded chitin fibers is primarily responsible for the mechanical behavior of the cuticle, although this is modulated by the degree of tanning and water retention. \cite{Klocke2011,Vincent2004}.	
		Accordingly, cuticle is often considered as a fiber-reinforced composite material; however, in beetles, the composite arrangement of the chitin fibers varies by cuticle region.

@@

	\section{Tensile testing of the rostrum} %shoot for 600 words
		To better characterize the mechanical behavior of the snout as an integrated whole, we performed tensile testing on the snouts of six species in the genus \textit{Curculio}, representing a mixture of closely and distantly related taxa.
		Although the heads were rehydrated by immersion in de-ionized water for 24 hours, we observed comparatively brittle fracture (but see below), in contrast to GIVE EXAMPLES HERE.
		The tensile behavior for the cuticle of these weevils is a result of its microstructure, which lacks pores, etc (give reasons).
		THIS IS ALSO A GOOD SPOT TO SIGNPOST WHAT I FOUND IN GENERAL.

		\subsection{Tensile behavior}
			We observed that the maximum force sustained at the site of the break was strongly correlated with the cross-sectional area of the endocuticle, and not the exocuticle.
			Consequently, there was a negative correlation between utlimate tensile strenth of the specimen and the ratio of exocuticle to endocuticle cross-sectional area at the site of fracture, indicating that as the proportion of exocuticle increased, tensile strength decreased.
			In other words, UTS is strongly correlated with the cross-sectional area of the endocuticle across species.
			These associations were found to be statistically significant and independent of species membership, rostrum length, and location on the snout.

@@

		\subsection{Fractography}
			Examination of the fracture surfaces and adjacent cuticle of tensile testing specimens revealed that although fracture was comparatively brittle, the fracture mechanics of the exocuticle and endocuticle differed according to their microstructure, in agreement with previous studies (name some).

@@

			From this pattern we hypothesize that the exocuticle-rich gular sutures are the most likely site for the initiation of void nucleation and failure of the integrated rostral cuticle in cross section.
			Structural failure would take place as cracks propagate through the endocuticle from these sutures, which penetrate the entire thickness of the laminate.
			We speculate that this could be the reason why the cross-sectional profile of \textit{C. caryae} is flattened ventrally
			Flattening this region may reduce tensile-strain across the gular sutures when the snout is bent dorsally, thus reducing the risk of fracture in the elongate, strongly-curved rostrum in this species.

	\section{Fatigue testing of \textit{Curculio caryae}} %shot for 600 words
		The shearing motion and accommodation of strain led us to question how the cuticle might accommodate repeated strain, as is seen in the living organism.
		We therefore performed fatigue testing on a female curculio longinasus, which exhibits the most extreme degree of bending in the sis species examined.

	\section{Conclusions} %shoot for 500 words
			Overall conclusion - the only difference between the flexible part and the stiff part is the microstructure; there is no difference in the mechanical behavior of the endocuticle at different locations on the snout and across species.

%actual methods after references
	\section{Methods}
		Methods, including statements of data availability and any associated accession codes and references, are available in the online version of this paper.

%50 references (as a guideline, in order of appearance in text)
	\begin{thebibliography}{50}
		\section*{References}	
			\bibitem{Klocke2011}
				Klocke, D. \& Schmitz, H.
				Water as a major modulator of the mechanical properties of insect cuticle.
				\textit{Acta Biomater.}
				\textbf{7,}
				2935--2942
				(2011).

>>>>

%TODO NOTES! REMOVE FROM COMPLETE DRAFT-----_-----------------------------------
\usepackage{xargs}
\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
\newcommandx{\unsure}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
\newcommandx{\change}[2][1=]{\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{#2}}
\newcommandx{\info}[2][1=]{\todo[linecolor=green,backgroundcolor=green!25,bordercolor=green,#1]{#2}}
\newcommandx{\improvement}[2][1=]{\todo[linecolor=purple,backgroundcolor=purple!25,bordercolor=purple,#1]{#2}}
\newcommandx{\thiswillnotshow}[2][1=]{\todo[disable,#1]{#2}}
%-------------------------------------------------------------------------------

@@

	%title and author block
	{\title{Exoskeletal microstructure and tensile behavior of the acorn weevil rostrum}
	%CONSIDER CHANGING THE TITLE AFTER WRITING MANUSCRIPT
	\date{\today}

@@

{%manuscript outline:
%3000 word limit, 150 word abstract, 500 word intro, 250 word fig caps

%standard intro, talk about behavior, rostrum bending, and cuticle
%weevil cuticle is known to be stronger in compression than in tension
%in Curculio this means that the greatest risk of breakage would come from tension during bending of the snout
%curculio appear to have mitigated the risk of breakage by making the rostral apex flexible
%prior work suggests that this flexibility comes from a modified composite profile in the cuticle of the rostral apex
%different parts of the head thus appear to have differently specialized cuticle to deal with extreme bending
%only observed in 1 species, never any comparison across species/in genus, fractographic analysis never attempted to explain how snout is most likely to break/ how the cuticle fails (ie what failure modes are the weevils fighting against/trying to mitigate)

%we therefore set out to characterize the modifications to the cuticle of the rostral apex that presumably allow bending, using six species representing a mix of phylogenetically related and disparate species.
%because material rigidity and toughness are often inversely proportional, we also investigate how the composite structure of the cuticle contributes to the tensile strength of the snout, whether the flexible regions exhibit higher tensile strenth, and conduct fractographic analysis on tensile testing specimens to determine the failure modes of the rostrum/cuticle under tension.
%We additionally present the results of a load cycling test for a representative specimen of a species with a highly elongate and curved snout to explore the possibility that bending of the rostrum results in elastic deformation of the cuticle.

%This study demonstrates: (1) there are 2 microstructural modifications to the composite profile of the cuticle of the rostral apex that would enhance cuticular flexibility, present in all examined species, and likely across the genus (2) thickness of the encocuticle is directly proportional to the maximum tensile force sustained in tension, to the exclusion of the exocuticle, (3) the tensile strenth of the snout varies across the length of the head and is inversely proportional to exo/endocuticle ratio, such that the more flexible areas are also stronger under tension, (4) young's modulus is higher for longer snouts, reflecting increased reinforcement at the base of the snout, likely to prevent bucking fracture at the base, (5) because of the modifications to the cuticle of the rostral apex (which result in all of these changes to the mechanical properties of the snout), bending appears to be viscoelastic/elastic, with a relatively low risk of breakage during the lifetime of the beetle, (6) when the rostrum does fail, crack initiation appears to take place in the exocuticle-rich gular sutures, which would be under tension during the drilling behavior in the live insect.

%in conclusion, the cuticle of the rostral apex is modified to increase flexibility, allowing for nearly elastic bending, while simultaneously increasing tensile strength, thus resisting the predominant fracture modes observed in unaltered cuticle.
%to our knowledge, this is the first time that modifications to the composite profile of the cuticle have been observed to optimized mechanical behavior in insects, or something like that.

%this has important implications for understanding the evolution of this highly exaggerated structure, ie suggesting negative selection of breakage instead of positive selection of flexibility. possibly a single ancestral shift for each modification, opened a large region of morphospace for exploration, resulting in present day diversity of the genus.

%-------------------------------------------------------------------------------
%introduction - nature materials requires that introduction is given NO SECTION HEADER, 500 words
}	
\section*{Introduction}
%Intro outline - 500 word maximum
%	Paragraph 1 - Cuticle of beetles
%	Paragraph 2 - Weevils and behavior
%	Paragraph 3 - modified cuticle in curculio
%	Paragraph 4 - study objectives
%	Paragraph 5 - short segue/ signposting
	{The exoskeleton of Coleoptera (beetles) is a hierarchically-structured fibrous composite typified by variously arranged $\alpha$-chitin (N-acetylglucosamine) nanofibrils embedded in a heterogeneous protein matrix.
	Although $\alpha$-chitin is brittle and strongly anisotropic\unsure{give some numbers here}, beetle cuticle is simultaneously rigid and tough due to its uniquely layered microstructure\unsure{give more numbers here}.
	Beetle cuticle is divided into two structurally and mechanically distinct regions, the (outer) exocuticle and (inner) endocuticle.
	The exocuticle is characterized by a transversely-isotropic helicoidal structure that is comparatively rigid (though brittle) and has been the subject of many prior studies\unsure{numbers and refs}.
	By contrast, the endocuticle of beetles is comprised of large, aligned bundles (macrofibers) of chitin that are strongly anisotropic\unsure{numbers again}.
	Arranged in unidirectional laminae, the macrofibers contribute to the toughness of the exoskeleton by inhibiting crack formation and propagation between successive plies \cite{Kamp2010,Kamp2015,Hepburn1973}.

	In general, impact-prone areas and exaggerated structures, such as horns and legs exhibit cuticle organization that resists deformation and fracture.
	Acorn weevils in the genus \textit{Curculio} are typified by an elongate structure of the head, called the rostrum (snout), which instead exhibits unusual distal flexibility.
	The rostrum is a hollow, strongly curved (over 90$\degree$ in some species), cylindrical, exoskeletal extension of the otherwise nearly-spherical head, which bears at its apex the terminal chewing mouthparts.
	Despite being composed of the same material as other rigid body parts, the snout can be repeatedly bent without evident damage.
	This structure is used by the female to feed and excavate sites for egg-laying (oviposition); the latter process causes significant, apparently elastic, deformation of the rostrum\improvement{as illustrated in figure 1c}.
	By maintaining constant tension on the snout and rotating around the bore-hole, females are able to flex the rostrum into a near-perfectly straight configuration and thereby produce a linear channel into the fruit.

	While this behavior has been observed in many species of \textit{Curculio}, it was unclear how the rostrum of female acorn weevils can withstand the repeated, often extreme bending incurred during the process of egg-chamber excavation.
	In this study we characterize the composite profile of the rostral cuticle to account for the observed flexibility of the snout.
	We show that the relative layer thicknesses and fiber orientation angles of the exocuticle and endocuticle of the rostrum are strongly differentiated from the head capsule and other body parts.
	Because recent studies have shown that the yield strength of the weevil rostrum exoskeleton is lower in tension than compression, we perform a comparative analysis of the ultimate tensile strength of the rostrum across species and snout morphotypes; we also report the results of displacement-controlled load cycling of the snout in a species with strongly curved morphology.
	We relate an observed increase in the volume fraction of endocuticle in the rostrum to higher tensile strength at the rostral apex in all tested species, and find that a strongly curved rostrum can be flexed repeatedly without harm to the structure.

	We additionally describe the fracture mechanics of the snout, as pertains to both cuticle composite structure and tensile behavior, and consider how modification of the cuticle may reduce the risk of rostral fracture during oviposition.
	Based on our findings we posit that the composite profile of the rostral apex enables the snout to be flexed until straight while remaining within the elastic limits of the material, mitigating the risk of structural damage, and without evident alteration of the mechanical properties of the individual components of the cuticle across the structure and between species.
	Thus, the flexibility and tensile strength of the rostrum appear to be derived \emph{exclusively} from modification of the composite architecture of the exoskeleton.
	To our knowledge, this is the first time that a modified composite profile has been reported as a means of enhancing structural elasticity in the insect exoskeleton.\info{630 words, reduce to 600 or less in revisions, structure of argument/narrative is good}
}	
	\section{Microstructure of the \textit{Curculio} rostrum}\label{sec:microstructure}
		Insect cuticle is made up of chitin and numerous uncharacterized proteins, which effectively act as a matrix into which the chitin is embedded \cite{Nikolov2011,Nikolov2010,Vincent2004}.
		The arrangement of the embedded chitin fibers is primarily responsible for the mechanical behavior of the cuticle, although this is modulated by the degree of tanning and water retention. \cite{Klocke2011,Vincent2004}.	
		Accordingly, cuticle is often considered as a fiber-reinforced composite material; however, in beetles, the composite arrangement of the chitin fibers varies by cuticle region.

@@

	\section{Force-controlled loading to fracture}
		To better characterize the mechanical behavior of the snout as an integrated whole, we performed tensile testing on the snouts of six species in the genus \textit{Curculio}, representing a mixture of closely and distantly related taxa.
		Although the heads were rehydrated by immersion in de-ionized water for 24 hours, we observed comparatively brittle fracture (but see below), in contrast to GIVE EXAMPLES HERE.
		The tensile behavior for the cuticle of these weevils is a result of its microstructure, which lacks pores, etc (give reasons).
		THIS IS ALSO A GOOD SPOT TO SIGNPOST WHAT I FOUND IN GENERAL.

		We observed that the maximum force sustained at the site of the break was strongly correlated with the cross-sectional area of the endocuticle, and not the exocuticle.
		Consequently, there was a negative correlation between utlimate tensile strenth of the specimen and the ratio of exocuticle to endocuticle cross-sectional area at the site of fracture, indicating that as the proportion of exocuticle increased, tensile strength decreased.
		In other words, UTS is strongly correlated with the cross-sectional area of the endocuticle across species.
		These associations were found to be statistically significant and independent of species membership, rostrum length, and location on the snout.

@@

	\section{Load cycling of \textit{Curculio caryae}}
		The shearing motion and accommodation of strain led us to question how the cuticle might accommodate repeated strain, as is seen in the living organism.
		We therefore performed fatigue testing on a female curculio longinasus, which exhibits the most extreme degree of bending in the sis species examined.

	\section{Fractography of test specimens}
		Examination of the fracture surfaces and adjacent cuticle of tensile testing specimens revealed that although fracture was comparatively brittle, the fracture mechanics of the exocuticle and endocuticle differed according to their microstructure, in agreement with previous studies (name some).

@@

		From this pattern we hypothesize that the exocuticle-rich gular sutures are the most likely site for the initiation of void nucleation and failure of the integrated rostral cuticle in cross section.
		Structural failure would take place as cracks propagate through the endocuticle from these sutures, which penetrate the entire thickness of the laminate.
		We speculate that this could be the reason why the cross-sectional profile of \textit{C. caryae} is flattened ventrally.
		Flattening this region may reduce tensile-strain across the gular sutures when the snout is bent dorsally, thus reducing the risk of fracture in the elongate, strongly-curved rostrum in this species.

	\section{Conclusions}
	%1-2 'short paragraphs'

	\section{Methods}
	%actual methods after references
		Methods, including statements of data availability and any associated accession codes and references, are available in the online version of this paper.

	\begin{thebibliography}{50}
	%50 references (as a guideline, in order of appearance in text)
		\section*{References}	
			\bibitem{Klocke2011}
				Klocke, D. \& Schmitz, H.
				Water as a major modulator of the mechanical properties of insect cuticle.
				\textit{Acta Biomater.}
				\textbf{7,}
				2935--2942
				(2011).
